% \begin{Dedication}
% \vfill
% \begin{flushright}
% \textit
% {
%     This thesis decicated to someone.
% }
% \end{flushright}
% \end{Dedication}


%\begin{Extra}
%\end{Extra}


\begin{Acknowledgments}
\setlength{\baselineskip}{1.5\baselineskip}
{

    First and foremost, thanks to my research advisor and mentor Martin Wosnik,
    without whose (in my opinion, far-fetched) confidence and unwaivering
    support none of this work would have even been undertaken. I'd also like to
    thank:

    \begin{itemize}

        \item My committee---professors Baldwin, Foster, Lippmann, Polagye, and
        Swift---for their expertise and guidance.

        \item Ashley, for trying to make sure I never worked too hard, and for
        understanding when I did.

        \item My parents for giving me a place to be creative growing up, my
        family for always asking ``so when are you going to be done with
        school?'', and Tom and Gramps for helping out the with tow tank.

        \item Paul Lavoie for many words of wisdom, and allowing for plenty of
        ``third shift'' time in the machine shop.

        \item Chris Ellis and Jim Mullin at SAFL for their advice regarding the
        tow tank linear guide system upgrades.

        \item Les Peabody, Craig Deady, Sean Mitchell from Minarik for their
        help with the upgraded tow tank motion system electronics.

        \item Lucid Energy Technologies, LLP, for materials used to build the
        UNH-RVAT turbine and experimental setup.

        \item All the open-source software projects, developers, and communities
        for sharing their skills and support: Python, Anaconda, NumPy,
        Matplotlib, Pandas, OpenFOAM (The OpenFOAM Foundation, Bruno Santos,
        Chris Sideroff, Vesselin Krastev), swak4Foam, Linux, Ubuntu, Linux Mint,
        Git, NREL (SOWFA), and more, used in this work.

        \item Anders Goude for contributions to the turbinesFoam actuator line
        model.

        \item Matt Rowell and Vincent Lyon for their help conducting the
        experiments.

    \end{itemize}

    Finally, I'd like to aknowledge financial support from the US Department of
    Energy, from Dr. Geoffrey A. Prentice and the National Science Foundation
    (award 1150797), and Dr. Vincent S. Neary and Sandia National Labs.

}
\end{Acknowledgments}
