\chapter*{Appendix A: Sharing research outputs}

It is unknown how many experimental campaigns have been redundant due to lack of
openness in the research process. While one may argue this is good for the
repeatability of research, it can lead to ``reinvention of the wheel,'' whose
inefficiency is especially important in this age of diminishing research
funding. It was therefore a secondary goal of all the work performed herein was
to share the products in an open and useful manner.

One important objective is to distribute research products (code, data, CAD
files, etc.) openly, such that they are as useful and reproducible as possible.
It is also important that the code is set up to grow and evolve via
contributions from the community---one of the most significant advantages of the
open-source development model. One example occurred during this research, where
an OpenFOAM case, hosted openly on GitHub, received contributions from a random
user, since he or she was going to create something similar, but now didn't have
to.

Unfortunately, a lot of research code is merely described in publications, but
never released for adoption and improvement by the community. For example,
Shamsoddin and Porte-Agel have written a cross-flow turbine actuator line model
for LES, but have not made their code available or described results regarding
performance prediction \cite{Shamsoddin2014}. Even CACTUS, despite being
open-source, is simply distributed as a static archive for download, so there is
no clear procedure for users to contribute to the code.

To help correct this issue, the results of this research will be open-source and
hosted on GitHub.com to facilitate the inclusion of improvements by other
developers through their fork/pull mechanism. We have already implemented a
sharing strategy for experimental results using GitHub to host code and
processed data, and figshare.com for raw data, CAD files, and snapshots of code
or simulation settings used in specific publications
\cite{Bachant2014-RVAT-baseline, Bachant2014-RVAT-CAD,
Bachant2014-OF-AS-case-files, Bachant2015-RVAT-Re-dep-data}. We will continue to
share results in this way to ensure maximum impact from reuse, remixing, etc.
Furthermore, the library developed here will be openly available through all
stages of development (in fact, it is already), which will hopefully prevent
other developers from unknowingly working on similar projects, and collaborating
instead.